---
title: "Integration"
author: "Youyung Lim"
date: "22/12/2021"
output: html_document
---

### Integration - Seurat Standard Workflow
https://satijalab.org/seurat/archive/v3.0/integration.html

Datasets are log-normalized, and the features that are repeatedly variable across datasets (anchors) are identified between all pairs (i.e. 10 datasets = 45 pairwise comparisons). While equal weights are given to the datasets, it is computationally intensive compared to the reference-based workflow which reduces the number of comparison by specifying one or more of the datasets as the ‘reference’ and the rest as queries. 

Different options* did not give different results in the integrated UMAP. Hence we stick to the standard workflow.
*
Log-normalization                   / Standard / nfeature = 1000, 2000, 3000
Log-normalization / Reciprocal PCA / Standard / nfeature = 1000, 2000, 3000
Log-normalization / Reciprocal PCA / Reference-based / nfeature = 2000

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(Seurat)
library(patchwork)
library(tidyr)
library(stats)

path <- '/mnt/resu0001/Boonstra/Users/Youkyung/healthyLiver'
path2 <- '/mnt/resu0001/Boonstra/Users/Youkyung/integrated'
```

Load the preprocessed datasets which is saved in a list
```{r}
liver.list0 <- readRDS(file = file.path(path2,'5/preprocessed_datasets.rds'))
```

### Integration
```{r}
liver.list <- lapply(X = liver.list0, FUN = function(x) {
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = liver.list, nfeatures = 2000)
liver.anchors <- FindIntegrationAnchors(object.list = liver.list, anchor.features = features)

# # Reciprocal PCA
# liver.list <- lapply(X = liver.list, FUN = function(x) {
#     x <- ScaleData(x, features = features, verbose = FALSE)
#     x <- RunPCA(x, features = features, verbose = FALSE)
# })
#
# liver.anchors <- FindIntegrationAnchors(object.list = liver.list, anchor.features = features, reduction = "rpca")
```

```{r}
# Create an 'integrated' data assay
liver.integrated <- IntegrateData(anchorset = liver.anchors)
DefaultAssay(liver.integrated) <- "integrated"
```


### Save the datasets as scHPL input format
Transponse and save normalized data '@data' as data.csv
Save cluster labels as labels.csv

```{r}
# Select 
macpa1 <- liver.integrated@meta.data$project == 'MacParland2018'
macpa2 <- liver.integrated@meta.data$project == 'MacParland2021'
tambu <- liver.integrated@meta.data$project == 'Tamburini'
aizar <- liver.integrated@meta.data$project == 'Aizarani'
segal <- liver.integrated@meta.data$project == 'Segal'

macparland1_label <- liver.integrated@meta.data$cluster[macpa1]
macparland1_integrated <- t(as.data.frame(liver.integrated@assays$integrated@data[,macpa1]))
macparland2_label <- liver.integrated@meta.data$cluster[macpa2]
macparland2_integrated <- t(as.data.frame(liver.integrated@assays$integrated@data[,macpa2]))
tamburini_label <- liver.integrated@meta.data$cluster[tambu]
tamburini_integrated <- t(as.data.frame(liver.integrated@assays$integrated@data[,tambu]))
aizarani_label <- liver.integrated@meta.data$cluster[aizar]
aizarani_integrated <- t(as.data.frame(liver.integrated@assays$integrated@data[,aizar]))
segal_label <- liver.integrated@meta.data$cluster[segal]
segal_integrated <- t(as.data.frame(liver.integrated@assays$integrated@data[,segal]))
```


```{r}
write.csv(macparland1_integrated, file = file.path(path2,'5/macparland2018_data.csv'), row.names = TRUE)
write.csv(macparland1_label, file = file.path(path2,'5/macparland2018_label.csv'), row.names = FALSE)
write.csv(macparland2_integrated, file = file.path(path2,'5/macparland2021_data.csv'), row.names = TRUE)
write.csv(macparland2_label, file = file.path(path2,'5/macparland2021_label.csv'), row.names = FALSE)
write.csv(tamburini_integrated, file = file.path(path2,'5/tamburini_data.csv'), row.names = TRUE)
write.csv(tamburini_label, file = file.path(path2,'5/tamburini_label.csv'), row.names = FALSE)
write.csv(aizarani_integrated, file = file.path(path2,'5/aizarani_data.csv'), row.names = TRUE)
write.csv(aizarani_label, file = file.path(path2,'5/aizarani_label.csv'), row.names = FALSE)
write.csv(segal_integrated, file = file.path(path2,'5/segal_data.csv'), row.names = TRUE)
write.csv(segal_label, file = file.path(path2,'5/segal_label.csv'), row.names = FALSE)
```


```{r}
# Save the integrated dataset with nfeature=2000
saveRDS(liver.integrated, file = file.path(path2,'5/integrated_2000.rds'))
```

 

### Integrated Analysis
Decide nPC using elbow plot and variance explained 
```{r}
# Elbowplot
liver.integrated <- ScaleData(liver.integrated, verbose = FALSE)
liver.integrated <- RunPCA(liver.integrated, npcs = 30, verbose = FALSE)
ElbowPlot(liver.integrated, ndims = 30)
png(filename = file.path(path2,'5/elbow.png'), width = 800, units = 'px')
```

```{r}
# Variance Explained (total sum of squares)
# ssqtotal = sum(X1*X1) where X1 is a value substracted column means
# SSQ = variance explained = 100* (1-sum((X1-Xest)^2)/ssqtotal)

Xm <- liver.integrated@assays$integrated@scale.data
total_variance <- sum(matrixStats::rowVars(Xm)) 
eigenValues <- (liver.integrated@reductions$pca@stdev)^2
varExplained <- 100 * (eigenValues / total_variance)

npc = 30
ssqtot <- 0 * (1:npc)
for (i in 1:npc){
  ssqtot[i] <- sum(varExplained[1:i])
}

write.csv(data.frame(npc=c(1:npc), ssq=varExplained, ssqtot=ssqtot), file = file.path(path2,'5/VarExplained3.csv'))
```


Generate UMAP plots (nPC = 15)
```{r}
liver.integrated <- ScaleData(liver.integrated, verbose = FALSE)
liver.integrated <- RunPCA(liver.integrated, npcs = 15, verbose = FALSE)
liver.integrated <- RunUMAP(liver.integrated, reduction = 'pca', dims = 1:15)
# liver.integrated <- RunTSNE(liver.integrated15, reduction = 'pca', dims = 1:15)
# liver.integrated <- FindNeighbors(liver.integrated, reduction = 'pca', dims = 1:15)
```

UMAP with colors of each dataset per a metadata (percent.mt, project, cluster, sample). 
No noticeable bias has been found in the integrated dataset.
```{r}
# UMAP with colors of each dataset per a metadata
DimPlot(liver.integrated, reduction = "umap", group.by = 'percent.mt') + NoLegend()
DimPlot(liver.integrated, reduction = "umap", group.by = 'project', label = TRUE, repel = TRUE)
DimPlot(liver.integrated, reduction = "umap", group.by = 'cluster', label = TRUE, repel = TRUE) + NoLegend()
DimPlot(liver.integrated, reduction = "umap", group.by = 'sample', label = TRUE, repel = TRUE) + NoLegend()
```

```{r}
# Save umap_project
p1 <- DimPlot(liver.integrated, reduction = "umap", group.by = 'project', label = FALSE, repel = TRUE, cols = c("#E16A86", "#909800", "#00AD9A", '#619CFF', '#9183E6'))
p1$data$project <- factor(x = p1$data$project, levels = c('MacParland2018', 'MacParland2021', 'Tamburini', 'Aizarani', 'Segal'))
png(filename = file.path(path2,'5/umap_project.png'), width = 800, units = 'px')
p1
```

To take a better look at the 'umap_project.png' figure, we will split this into five.
UMAP of each dataset with its labels - MacParland 2018
```{r}
# Get the cells for each dataset (per project)
Idents(liver.integrated) <- liver.integrated$project
MacParland2018 <- WhichCells(liver.integrated, idents = "MacParland2018")
MacParland2021 <- WhichCells(liver.integrated, idents = "MacParland2021")
Tamburini <- WhichCells(liver.integrated, idents = "Tamburini")
Aizarani <- WhichCells(liver.integrated, idents = "Aizarani")
Segal <- WhichCells(liver.integrated, idents = "Segal")

# Remove original cluster information, and add back the cluster information from only one of the datasets
liver.integrated1 <- liver.integrated
liver.integrated1$cluster <- NA
liver.integrated1$cluster <- liver.integrated[,liver.integrated$project =="MacParland2018"]$cluster

# Cells with NA in $cluster will be shown in grey, and the cells from the one dataset will be highlighted.
png(filename = file.path(path2,'5/umap_macparland2018.png'), width = 800, units = 'px')
DimPlot(liver.integrated1, label = T, repel = T, group.by = 'cluster', label.size = 4, reduction = "umap", cells.highlight= MacParland2018, cols.highlight = "#E16A86", cols= "grey") + NoLegend()
dev.off()
```

UMAP of each dataset with its labels - MacParland 2021
```{r}
liver.integrated1$cluster <- NA
liver.integrated1$cluster <- liver.integrated[,liver.integrated$project =="MacParland2021"]$cluster
png(filename = file.path(path2,'5/umap_macparland2021.png'), width = 800, units = 'px')
DimPlot(liver.integrated1, label = T, repel = T, group.by = 'cluster', label.size = 4, reduction = "umap", cells.highlight= MacParland2021, cols.highlight = "#909800", cols= "grey")+ NoLegend()
dev.off()
```

UMAP of each dataset with its labels - Tamburini
```{r}
liver.integrated1$cluster <- NA
liver.integrated1$cluster <- liver.integrated[,liver.integrated$project =="Tamburini"]$cluster
png(filename = file.path(path2,'5/umap_tamburini.png'), width = 800, units = 'px')
DimPlot(liver.integrated1, label = T, repel = T, group.by = 'cluster', label.size = 5, reduction = "umap", cells.highlight= Tamburini, cols.highlight = "#00AD9A", cols= "grey")+ NoLegend()
dev.off()
```

UMAP of each dataset with its labels - Aizarani
```{r}
liver.integrated1$cluster <- NA
liver.integrated1$cluster <- liver.integrated[,liver.integrated$project =="Aizarani"]$cluster
png(filename = file.path(path2,'5/umap_aizarani.png'), width = 800, units = 'px')
DimPlot(liver.integrated1, label = T, repel = T, group.by = 'cluster', label.size = 5, reduction = "umap", cells.highlight= Aizarani, cols.highlight = "#619CFF", cols= "grey")+ NoLegend()  
dev.off()
```

UMAP of each dataset with its labels - Segal
```{r}
liver.integrated1$cluster <- NA
liver.integrated1$cluster <- liver.integrated[,liver.integrated$project =="Segal"]$cluster
png(filename = file.path(path2,'5/umap_segal.png'), width = 800, units = 'px')
DimPlot(liver.integrated1, label = T, repel = T, group.by = 'cluster', label.size = 5, reduction = "umap", cells.highlight= Segal, cols.highlight ="#9183E6", cols= "grey")+ NoLegend()
dev.off()
```


Checking B cells after removing a cluster 'Bcell_Mature B cells'. It has been removed due to the following reasons.
• The cluster has mere 15 cells which is only slightly above the threshold of removing a cluster with < 9 cells.
• It is not properly clustered / scattered.
• It could not be integrated well with other B cell clusters.
```{r}
# B cell clusters after removing "Bcell_Mature B cells" from MacParland2021
bcells <- colnames(liver.integrated[,liver.integrated$cluster %in% c("Mature_B_Cells", "Plasma_Cells",  "Lymph_Mature B cells", "Lymph_Plasma B cells", "Resident B cells", "MHC II+ B cells")])

liver.integrated1 <- liver.integrated
liver.integrated1$cluster <- NA
liver.integrated1$cluster <- liver.integrated[,liver.integrated$cluster %in% c("Mature_B_Cells", "Plasma_Cells", "Lymph_Mature B cells", "Lymph_Plasma B cells", "Resident B cells", "MHC II+ B cells")]$cluster

png(filename = file.path(path2,'5/umap_bcells.png'), width = 800, units = 'px')
DimPlot(liver.integrated1, label = T, repel = T, group.by = 'cluster', label.size = 4, reduction = "umap", cells.highlight = bcells, cols.highlight ="#FFCC00", cols= "grey") + NoLegend() + ggtitle("B cell clusters")
dev.off()
```
